Ionospheric Disturbances in Mexican Territory Produced by Objects Entering the Atmosphere from Space \tnoteref{tnote1}

### 2021-11-16 paper/methodology.tex
Emergency Backup

\begin{table*}
  \centering
  \caption{List of meteors passing through Mexico. The events are listed in chronological order. The listed duration, latitude and longitude correspond to the mean of the measurements of both GOES satellites. The uncertainties correspond to the respecting mean deviation.}
\label{tab:table-meteors}
\begin{tabular}{rrrrrrr}
\hline
ID & Date of event & Start Time (UT)  & Duration (seconds) & Latitude (deg) & Longitude (deg) & Height (km)\\
\hline
01 & 2019-05-23 & 16:36:18 & $0.197\pm 0.0000$ & $24.30 \pm 0.000$ & $-101.60 \pm 0.849$  & 28\\
02 & 2019-07-18 & 14:30:30 & $0.058\pm 0.0000$ & $27.20 \pm 0.000$ & $-103.15 \pm 0.778$  & 72\\
03 & 2019-08-10 & 11:18:48 & $0.199\pm 0.0757$ & $21.50 \pm 0.000$ & $-102.50  \pm 0.849$ & 92\\
04 & 2019-10-03 & 07:55:33 & $0.106\pm 0.0297$ & $25.65 \pm 0.071$ & $-96.25 \pm   0.778$ & 74\\
05 & 2019-10-09 & 06:08:11 & $0.103\pm 0.0078$ & $23.60 \pm 0.000$ & $-111.95 \pm  0.212$ & 32\\
06 & 2019-11-16 & 09:36:04 & $0.396\pm 0.0134$ & $20.30 \pm 0.000$ & $-100.55 \pm  0.919$ & 82\\
07 & 2019-11-17 & 15:36:01 & $0.116\pm 0.0035$ & $31.70 \pm 0.000$ & $-117.70 \pm  1.131$ & 88\\
08 & 2019-11-19 & 07:57:40 & $0.097\pm 0.1138$ & $20.00 \pm 0.000$ & $-88.40 \pm  1.131$  & 99\\
09 & 2019-11-26 & 13:23:20 & $0.078\pm 0.0290$ & $23.90 \pm 0.000$ & $-108.70 \pm  0.849$ & 81\\
10 & 2019-12-04 & 09:42:54 & $0.173\pm 0.0028$ & $31.50 \pm 0.000$ & $-113.65 \pm  0.919$ & 77\\
11 & 2019-12-15 & 14:50:49 & $0.127\pm 0.0134$ & $27.70 \pm 0.000$ & $-114.10 \pm  0.849$ & 78\\
12 & 2019-12-29 & 16:16:35 & $0.062\pm 0.0134$ & $29.60 \pm 0.000$ & $-116.35 \pm  0.919$ & 79\\
13 & 2020-01-03 & 14:10:17 & $0.113\pm 0.0085$ & $30.20 \pm 0.000$ & $-117.65 \pm  0.919$ & 74\\
14 & 2020-01-06 & 16:39:27 & $0.118\pm 0.0042$ & $31.40 \pm 0.000$ & $-108.20 \pm  0.990$ & 81\\
15 & 2020-01-15 & 15:00:33 & $0.213\pm 0.1351$ & $19.45 \pm 0.071$ & $-95.55 \pm   0.919$ & 93\\
16 & 2020-02-12 & 09:25:40 & $0.210\pm 0.0226$ & $18.90 \pm 0.000$ & $-93.50 \pm   0.849$ & 90\\
17 & 2020-03-03 & 12:33:27 & $0.062\pm 0.0007$ & $18.25 \pm 0.071$ & $-106.35 \pm 0.636$  & 77\\
18 & 2020-03-31 & 19:31:52 & $0.105\pm 0.0573$ & $28.45 \pm 0.071$ & $-112.05 \pm  0.636$ & 61\\
19 & 2020-04-08 & 16:25:28 & $0.120\pm 0.0926$ & $26.10 \pm 0.000$ & $-93.90 \pm   0.849$ & 78\\
20 & 2020-04-18 & 17:43:25 & $0.139\pm 0.0106$ & $29.00 \pm 0.000$ & $-106.55 \pm  0.919$ & 82\\
21 & 2020-04-20 & 16:05:22 & $0.318\pm 0.1655$ & $28.15 \pm 0.071$ & $-97.85 \pm   1.061$ & 88\\
22 & 2020-04-25 & 11:03:09 & $0.323\pm 0.0813$ & $32.15 \pm 0.071$ & $-111.60 \pm  1.131$ & 84\\
23 & 2020-04-28 & 19:31:52 & $0.105\pm 0.0573$ & $28.45 \pm 0.071$ & $-112.05 \pm  0.636$ & 29\\
24 & 2020-05-08 & 10:06:16 & $0.490\pm 0.0750$ & $21.60 \pm 0.000$ & $-92.40 \pm   0.849$ & 81\\
25 & 2020-07-15 & 19:58:28 & $0.693\pm 0.0495$ & $24.00 \pm 0.000$ & $-108.35 \pm  0.495$ & 53\\
26 & 2020-08-07 & 13:29:57 & $0.163\pm 0.0057$ & $28.80 \pm 0.000$ & $-106.05 \pm  0.919$ & 89\\
27 & 2020-09-13 & 16:41:59 & $0.184\pm 0.0078$ & $28.45 \pm 0.071$ & $-113.75 \pm  0.919$ & 85\\
28 & 2020-09-30 & 12:28:11 & $0.100\pm 0.0078$ & $24.90 \pm 0.000$ & $-110.90 \pm  0.849$ & 83\\
29 & 2020-11-16 & 12:28:11 & $0.100\pm 0.0078$ & $24.90 \pm 0.000$ & $-110.90 \pm  0.849$ & 06\\
30 & 2020-11-17 & 12:53:41 & $0.404\pm 0.0262$ & $23.00 \pm 0.000$ & $-102.45 \pm  0.919$ & 93\\
31 & 2020-12-19 & 10:18:14 & $0.407\pm 0.0110$ & $21.95 \pm 0.071$ & $-101.60 \pm  0.990$ & 98\\
32 & 2020-12-23 & 09:43:01 & $0.148\pm 0.0014$ & $25.75 \pm 0.071$ & $-111.25 \pm  0.778$ & 81\\
33 & 2020-12-29 & 15:20:54 & $0.118\pm 0.0014$ & $16.80 \pm 0.000$ & $-102.20 \pm  0.707$ & 81\\
34 & 2021-03-31 & 09:01:17 & $0.753\pm 0.3083$ & $20.15 \pm 0.071$ & $-92.95 \pm  0.212$  & 24\\\hline
\end{tabular}
\end{table*}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{../meteors_map}
  \caption{Positions of events from table \ref{tab:table-meteors}. The label of each point correspond to the ID (first column) of the referred table.}
  \label{fig:meteors-map}
\end{figure}

\subsection{GPS data}

This material is based on services provided by the GAGE Facility, operated by UNAVCO, Inc., with support from the National Science Foundation and the National Aeronautics and Space Administration under NSF Cooperative Agreement EAR-1724794.

We got RINEX data from 3 to 7 stations depending of the event location and data availability that surround the event place in all directions as possible. A list of the stations where we got RINEX data is available in table  . Most of the stations lie in mexican territory, but in some cases we required data from other stations to cover events near the mexican frontier at north or south.

\clearpage
\onecolumn
\begin{landscape}
%\begin{table*}
%    \centering
  \begin{longtable}{lllp{4cm}p{10cm}}
      \caption{List of GPS stations used for this work.}
      \label{tab:table-stations}
      \endfirsthead
      \endhead

@@

    Station name & Latitude & Longitude & Date of events  & Citation \\\hline
    BAR1\hyperlink{Hudnut}{${}^1$}\hyperlink{Hudnut2}{${}^5$}& 33.48 & -119.03 & 2019-12-29 2020-01-03 & UNAVCO Community, Hudnut, Kenneth, King, Nancy, Aspiotes, Aris G., Borsa, Adrian A., Determan, Daniel N., Galetzka, John E., Stark, Keith F., 2005, SCIGN-PBO Nucleus GPS Network - BAR1-Santa Barbara Island One P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5668BHN}.\\
    BLYT\hyperlink{Hudnut}{${}^1$} & 33.61 & -114.71 & 2019-12-29 2020-01-03 & Hudnut, Kenneth, King, Nancy, Aspiotes, Aris G., Borsa, Adrian A., Determan, Daniel N., Galetzka, John E., Stark, Keith F., 2006, SCIGN USGS GPS Network - BLYT-Blythe P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5HT2MKK}.\\
    CN23 & 17.26 & -88.78 & 2019-11-19 2020-01-15 2020-02-12& UNAVCO Community, 2012, COCONet GPS Network - CN23-BelmopanBZCR2012 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5Q23XJH}.\\
    CN25 & 16.23 & -92.13 & 2020-01-15 & UNAVCO Community, 2014, COCONet GPS Network - CN25-ComitandDMEX2012 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T57W69G7}.\\
    GCFS & 19.31 & -81.18 & 2019-11-19& Watts, Anthony, 2016, COCONet GPS Network - GCFS-G\_CAYMAN\_CYM2014 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/7ETV-X536}.\\
    GMPK\hyperlink{Hudnut}{${}^1$} & 33.05 & -114.83 & 2019-12-04 & UNAVCO Community, Hudnut, Kenneth, King, Nancy, Aspiotes, Aris G., Borsa, Adrian A., Determan, Daniel N., Galetzka, John E., Stark, Keith F., 2005, SCIGN-PBO Nucleus GPS Network - GMPK-Glamis Peak P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/WCHN-H687}.\\
    GUAT\hyperlink{Garnier}{${}^2$} & 14.59 & -90.52 & 2020-02-12 & DeMets, Charles, Cosenza-Muralles, Beatriz, 2021, Central America 2018 - Guatemala, The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/KH2R-K704}.\\
    GUAX\hyperlink{Hudnut}{${}^1$} & 28.88 & -118.29 & 2019-10-09 2019-12-15 2019-12-29 2020-01-03 2020-03-31 2020-07-15 2020-09-13 2020-09-30 2020-12-23 & Hudnut, Kenneth, King, Nancy, Aspiotes, Aris G., Borsa, Adrian A., Determan, Daniel N., Galetzka, John E., Stark, Keith F., 2001, SCIGN USGS GPS Network - GUAX-Isla Guadalupe P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5GX48T2}.\\
    IAGX & 29.03 & -113.17 & 2019-12-04 & Gonzalez-Ortega, Alejandro, Galetzka, John E., Gonzalez, Javier, 2018, CICESE REGNOM GPS Network - IAGX-iagxREGNOMmx2018 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/DGWN-A627}.\\
    INEG & 21.85 & -102.28 & 2020-07-15 2020-08-07 2020-09-30 2020-11-16 2020-11-17 2020-12-19 & No citations were found\\
    KVTX & 27.55 & -97.89 & 2019-05-23 2019-07-18 2019-08-10 2019-10-03 2019-11-17 2020-04-08 2020-04-18 2020-04-20 2020-05-08 2020-08-07 & UNAVCO Community, 2007, PBO GPS Network - KVTX-KingsvilleTX2006 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5J38QH8}.\\
    MDO1 & 30.68 & -104.02 & 2019-07-18& No citations were found\\
    MGO5 & 30.68 & -104.02 & 2020-04-20 2020-08-07 & No citations were found\\
    MGW3 & 29.62 & -89.95 & 2020-04-08 2020-04-20 2020-05-08 & No citations were found\\
    OXTH & 16.29 & -95.24 & 2020-01-15 2020-02-12 & DeMets, Charles, Cabral-Cano, Enrique, 2008, Oaxaca GPS Network - OXTH-Tehuantepec P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5Q81B5V}.\\
    OXUM\hyperlink{Graham}{${}^3$} & 15.66 & -96.50 & 2021-03-31 & Cabral-Cano, Enrique, Salazar-Tlaczani, Luis, 2015, TLALOCNet - OXUM-oxum\_tnet\_mx2001 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5J964RP}.\\
    P001 & 31.95 & -112.80 & 2020-04 25 & UNAVCO Community, 2008, PBO GPS Network - P001-Organ\_PipeAZ2007 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5DR2SGP}.\\
    P014 & 31.97 & -11.09 & 2019-12-04 2019-12-29 2020-01-03 2020-01-06 2020-04-25 & UNAVCO Community, 2008, PBO GPS Network - P014-Sahuarita\_AZ2007 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5DJ5CMK}.\\
    P807 & 30.49 & -98.82 & 2019-11-17 2020-01-06 2020-04-20 2020-11-17 & UNAVCO Community, 2012, PBO GPS Network - P807-EcRockStPkTX2012 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5TQ5ZKM}.\\
    PLPX & 31.59 & -115.15 & 2019-12-04 & UNAVCO Community, 2011, PBO GPS Network - PLPX-Las\_PintasMX2010 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5K64G3T}.\\
    PTEX & 32.29 & -116.52 & 2019-12-29 2020-01-03 2020-09-13 2020-12-23 & UNAVCO Community, 2011, PBO GPS Network - PTEX-Testerazo\_MX2011 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5610XBP}.\\
    RG06 & 32.63 & -107.86 & 2020-04-25 & Sheehan, Anne, 2007, Rio Grande Rift GPS Network - RG06-RG06FaywodNM2006 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5668BFR}.\\
    RG07 & 32.50 & -106.84 & 2020-01-06 & Sheehan, Anne, 2007, Rio Grande Rift GPS Network - RG07-RG07CrucesNM2006 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5KD1W45}. \\
    SG33 & 31.77 & -106.51 & 2019-11-17 2020-04-18 2020-08-07 & Harder, Steven, Kaip, Galen, Montana, Carlos, 2004, SuomiNet-G GPS Network - SG33-UTEP P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T50863KQ}.\\
    TGMX  & 20.87 & -86.87 & 2021-03-31 & UNAVCO Community, 2015, COCONet GPS Network - TGMX-PtoMor\_TG\_MX2015 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5154FB7}.\\
    TNAM & 20.54 & -103.97 & 2020-03-03 2020-07-15 2020-09-30 2020-11-16 2020-11-17 2020-12-19 & UNAVCO Community, 2014, TLALOCNet - TNAM-TNAM\_TNET\_MX2014 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5QF8R4R}.\\
    TNAT & 18.13 & -98.04 & 2020-01-15 & UNAVCO Community, 2014, TLALOCNet - TNAT-TNAT\_TNET\_MX2014 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5G15Z4S}.\\
    TNBA & 28.97 & -113.55 & 2019-10-09 2019-11-26 2019-12-15 2019-12-29 2020-01-03& UNAVCO Community, 2015, TLALOCNet - TNBA-TNBA\_TNET\_MX2014 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T57M0688}.\\
    TNCC & 18.79 & -103.17 & 2020-03-03 & UNAVCO Community, 2015, TLALOCNet - TNCC-TNCC\_TNET\_MX2015 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T50R9MSK}.\\
    TNCM & 19.50 & -105.04 & 2020-03-03 2020-04-28 & UNAVCO Community, 2014, TLALOCNet - TNCM-TNCM\_TNET\_MX2014 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5B856FW}.\\
    TNCN & 18.55 & -101.97 & 2020-11-16 2020-12-29 & UNAVCO Community, 2016, TLALOCNet - TNCN-TNCN\_TNET\_MX2016 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5610XQM}.\\
    TNCU & 28.45 & -106.79 & 2019-05-23 2019-07-18 2019-08-10 2019-11-17 2019-12-15 2020-01-06 2020-03-31 2020-04-18 2020-07-15 2020-08-07 2020-11-17 2020-12-19 & UNAVCO Community, 2014, TLALOCNet - TNCU-CuauhtemocTN2014 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5V69GV2}.\\
    TNGF & 19.33 & -99.18 & 2020-11-16 2020-12-29 & Cabral-Cano, Enrique, Salazar-Tlaczani, Luis, 2016, TLALOCNet GPS Network - TNGF\_Geofisica-UNAM\_Mexico\_City\_TNET\_mx2015 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T53X851M}.\\
    TNHM & 29.08 & -110.97 & 2019-10-09 2019-11-26 2019-12-04 2019-12-15 2019-12-29 2020-01-03 2020-03-31 2020-04-18 2020-07-15 2020-08-07 2020-09-13 2020-09-30 2020-12-23 & UNAVCO Community, 2014, TLALOCNet - TNHM-hermosilloTN2014 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5KP80FV}.\\
    TNMS & 20.53 & -104.80 & 2019-10-09 2019-11-26 2019-12-15 2020-03-03 2020-07-15 & UNAVCO Community, 2014, TLALOCNet - TNMS-TNMS\_TNET\_MX2014 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T56H4FQ5}.\\
    TNNP & 16.12 & -97.14 & 2020-04-28 & Cabral-Cano, Enrique, Salazar-Tlaczani, Luis, DeMets, Charles, 2016, TLALOCNet - TNNP-tnnp\_tnet\_mx2015 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5N29V96}.\\
    TNNX & 17.41 & -97.22 & 2020-01-15 2020-02-12 2020-12-29 2021-03-31 & UNAVCO Community, 2014, TLALOCNet - TNNX-TNNX\_TNET\_MX2014 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T52R3PZ0}.\\

>>>>

\begin{table*}
  \centering
  \caption{List of bolides detected in mexican territory (plus one detected near Venezuela and one detected near Cuba), detected by the Geostationary Lighning mapper. The events are listed in chronological order. The listed duration, latitude and longitude correspond to the mean of the measurements of both GOES satellites. The uncertainties correspond to the respecting mean deviation.}
\label{tab:table-meteors}
\begin{tabular}{rrrrrrrrrr}

@@

ID & Date of event  & Start Time (UT)  & Duration (seconds) & Latitude (deg) & Longitude (deg)& Altitude (km) & Energy (kT )\\\hline
GLM-00 & 2019-02-01 & 18:17:09 & $2.651\pm 0.4907$ & $22.45 \pm 0.071$ & $-83.50  \pm  0.424$ & 24           & 2.0978196 +/- 0.42361329      \\         
GLM-01 & 2019-05-23 & 16:36:18 & $0.197\pm 0.0000$ & $24.30 \pm 0.000$ & $-101.60 \pm  0.849$ & 28           & 0.011954659 +/- 7.4054404e-3  \\
GLM-02 & 2019-07-18 & 14:30:30 & $0.058\pm 0.0000$ & $27.20 \pm 0.000$ & $-103.15 \pm  0.778$ & 72           & 5.7827709e-3 +/- 3.9334336e-3 \\
GLM-03 & 2019-08-10 & 11:18:48 & $0.199\pm 0.0757$ & $21.50 \pm 0.000$ & $-102.50  \pm 0.849$ & 92           & 0.010556123 +/- 6.6488245e-3  \\
GLM-04 & 2019-10-03 & 07:55:33 & $0.106\pm 0.0297$ & $25.65 \pm 0.071$ & $-96.25 \pm   0.778$ & 74           & 2.9915536e-3 +/- 2.2000998e-3 \\
GLM-05 & 2019-10-09 & 06:08:11 & $0.103\pm 0.0078$ & $23.60 \pm 0.000$ & $-111.95 \pm  0.212$ & 32           & 0.021837042 +/- 0.012429732   \\
GLM-06 & 2019-11-16 & 09:36:04 & $0.396\pm 0.0134$ & $20.30 \pm 0.000$ & $-100.55 \pm  0.919$ & 82           & 7.5423706e-3 +/- 4.9626060e-3 \\
GLM-07 & 2019-11-17 & 15:36:01 & $0.116\pm 0.0035$ & $31.70 \pm 0.000$ & $-117.70 \pm  1.131$ & 88           & 0.022397444 +/- 0.012701445   \\
GLM-08 & 2019-11-19 & 07:57:40 & $0.097\pm 0.1138$ & $20.00 \pm 0.000$ & $-88.40 \pm   1.131$ & 99           & 1.5012507e-3 +/- 1.1909667e-3 \\
GLM-09 & 2019-11-26 & 13:23:20 & $0.078\pm 0.0290$ & $23.90 \pm 0.000$ & $-108.70 \pm  0.849$ & 81           & 4.9551290e-3 +/- 3.4345742e-3 \\
GLM-10 & 2019-12-04 & 09:42:54 & $0.173\pm 0.0028$ & $31.50 \pm 0.000$ & $-113.65 \pm  0.919$ & 77           & 0.029149047 +/- 0.015891049   \\
GLM-11 & 2019-12-15 & 14:50:49 & $0.127\pm 0.0134$ & $27.70 \pm 0.000$ & $-114.10 \pm  0.849$ & 78           & 0.010556123 +/- 6.6488245e-3  \\
GLM-12 & 2019-12-29 & 16:16:35 & $0.062\pm 0.0134$ & $29.60 \pm 0.000$ & $-116.35 \pm  0.919$ & 79           & 4.2084911e-3 +/- 2.9746446e-3 \\
GLM-13 & 2020-01-03 & 14:10:17 & $0.113\pm 0.0085$ & $30.20 \pm 0.000$ & $-117.65 \pm  0.919$ & 74           & 0.011607116 +/- 7.2187549e-3  \\
GLM-14 & 2020-01-06 & 16:39:27 & $0.118\pm 0.0042$ & $31.40 \pm 0.000$ & $-108.20 \pm  0.990$ & 81           & 0.015448801 +/- 9.2392402e-3  \\
GLM-15 & 2020-01-15 & 15:00:33 & $0.213\pm 0.1351$ & $19.45 \pm 0.071$ & $-95.55 \pm   0.919$ & 93           & 0.012559739 +/- 7.7284725e-3  \\
GLM-16 & 2020-02-12 & 09:25:40 & $0.210\pm 0.0226$ & $18.90 \pm 0.000$ & $-93.50 \pm   0.849$ & 90           & 8.2107211e-3 +/- 5.3440405e-3 \\
GLM-17 & 2020-03-03 & 12:33:27 & $0.062\pm 0.0007$ & $18.25 \pm 0.071$ & $-106.35 \pm  0.636$ & 77           & 3.4441157e-3 +/- 2.4922397e-3 \\
GLM-18 & 2020-03-31 & 19:31:52 & $0.105\pm 0.0573$ & $28.45 \pm 0.071$ & $-112.05 \pm  0.636$ & 61           & 7.2469897e-3 +/- 4.7924800e-3 \\
GLM-19 & 2020-04-08 & 16:25:28 & $0.120\pm 0.0926$ & $26.10 \pm 0.000$ & $-93.90 \pm   0.849$ & 78           & 4.0292119e-3 +/- 2.8626279e-3 \\
GLM-20 & 2020-04-18 & 17:43:25 & $0.139\pm 0.0106$ & $29.00 \pm 0.000$ & $-106.55 \pm  0.919$ & 82           & 5.8303967e-3 +/- 3.9618242e-3 \\
GLM-21 & 2020-04-20 & 16:05:22 & $0.318\pm 0.1655$ & $28.15 \pm 0.071$ & $-97.85 \pm   1.061$ & 88           & 0.031378060 +/- 0.016913982   \\
GLM-22 & 2020-04-25 & 11:03:09 & $0.323\pm 0.0813$ & $32.15 \pm 0.071$ & $-111.60 \pm  1.131$ & 84           & 0.021997346 +/- 0.012507576   \\
GLM-23 & 2020-04-28 & 19:31:52 & $0.105\pm 0.0573$ & $28.45 \pm 0.071$ & $-112.05 \pm  0.636$ & 29           & 0.12687833 +/- 0.053736010    \\
GLM-24 & 2020-05-08 & 10:06:16 & $0.490\pm 0.0750$ & $21.60 \pm 0.000$ & $-92.40 \pm   0.849$ & 81           & 0.033207253 +/- 0.017743607   \\
GLM-25 & 2020-07-15 & 19:58:28 & $0.693\pm 0.0495$ & $24.00 \pm 0.000$ & $-108.35 \pm  0.495$ & 53           & 0.020548979 +/- 0.011800641   \\
GLM-26 & 2020-08-07 & 13:29:57 & $0.163\pm 0.0057$ & $28.80 \pm 0.000$ & $-106.05 \pm  0.919$ & 89           & 0.014606987 +/- 8.8040841e-3  \\
GLM-27 & 2020-09-13 & 16:41:59 & $0.184\pm 0.0078$ & $28.45 \pm 0.071$ & $-113.75 \pm  0.919$ & 85           & 0.010995616 +/- 6.8881646e-3  \\
GLM-28 & 2020-09-30 & 12:28:11 & $0.100\pm 0.0078$ & $24.90 \pm 0.000$ & $-110.90 \pm  0.849$ & 83           & 0.014013965 +/- 8.4951275e-3  \\
GLM-29 & 2020-11-16 & 12:28:11 & $0.100\pm 0.0078$ & $24.90 \pm 0.000$ & $-110.90 \pm  0.849$ &106           & 0.052044572 +/- 0.025868548   \\
GLM-30 & 2020-11-17 & 12:53:41 & $0.404\pm 0.0262$ & $23.00 \pm 0.000$ & $-102.45 \pm  0.919$ & 93           & 0.060624247 +/- 0.029367156   \\
GLM-31 & 2020-12-19 & 10:18:14 & $0.407\pm 0.0110$ & $21.95 \pm 0.071$ & $-101.60 \pm  0.990$ & 98           & 0.060272985 +/- 0.029225978   \\
GLM-32 & 2020-12-23 & 09:43:01 & $0.148\pm 0.0014$ & $25.75 \pm 0.071$ & $-111.25 \pm  0.778$ & 81           & 0.012127937 +/- 7.4982028e-3  \\
GLM-33 & 2020-12-29 & 15:20:54 & $0.118\pm 0.0014$ & $16.80 \pm 0.000$ & $-102.20 \pm  0.707$ & 81           & 0.013161054 +/- 8.0470909e-3  \\
GLM-34 & 2021-03-31 & 09:01:17 & $0.753\pm 0.3083$ & $20.15 \pm 0.071$ & $-92.95 \pm  0.212$  & 24           & 0.054259119 +/- 0.026782008   \\
GLM-Ven& 2019-06-22 & 21:25:45 & $4.873\pm 0.0000$ & $14.9 \pm 0.000$  & $-65.8 \pm 0.000$    & 25           & 6.1014359 +/- 0.81239700      \\ \hline
\end{tabular}
\end{table*}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{../meteors_map}
  \caption{Positions of events from table \ref{tab:table-meteors} (blue) and table \ref{tab:table-meteors-2}. The events GLM-00/USG-08 actually correpond to the same bolide, but there are little discrepancies about the position where the bolide was detecte. The same applies for the events GLM-23/USG-10 and GLM-Ven/USG-09, which not appears in the map.}
  \label{fig:meteors-map}
\end{figure}

By the other hand, we got another sample of 10 bolidesfrom US Goverment (USG) sensors from the Center for Near Earth Object Studies (CNEOS), publicly available at \url{https://cneos.jpl.nasa.gov/fireballs/}, where we may obtain directly data about bolides position, the date and time each bolide was detected, the energy released at fragmentation, the velocity and the height (the last two not availble for all bolides). As seen in table \ref{tab:table-meteors-2}, the time span is quite larger, and the released energy is generally larger. Both energy distributions are compared directly in figure \ref{fig:boxplot}. Some elements appear in both samples, since are bright enough to be detected regardless the project involved. In USG sample, the total energy of each meteor is obtained directly, but is not the case for the GLM sample. In appendix \ref{app:distance} and \ref{app:energy} we give details about how this total energy is obtained.

\begin{table*}
    \centering
  \footnotesize
  \caption{List of bolides detected in mexican territory (plus one detected near Venezuela and one detected near Cuba), detected by USG sensors.}
\label{tab:table-meteors-2}
\begin{tabular}{rrrrrrrrrrr}
       &               &                  & \multicolumn{4}{c}{Velocity (km/s)}       &                   &                 &\\\cline{4-7}
\multicolumn{1}{c}{ID}&\multicolumn{1}{c}{Date of event}&\multicolumn{1}{c}{Start Time (UT)}&\multicolumn{1}{c}{$v$}&\multicolumn{1}{c}{$v_x$}&\multicolumn{1}{c}{$v_y$}&\multicolumn{1}{c}{$v_z$}&\multicolumn{1}{c}{Latitude (deg)}&\multicolumn{1}{c}{Longitude (deg)}&\multicolumn{1}{c}{Altitude (km)} & \multicolumn{1}{c}{Energy (kT)}\\\hline
USG-01 & 1995-08-05    & 17:14:10         &      &         &           &              & 11.6              &  -104.3         &             &  0.56\\
USG-02 & 1996-07-12    & 14:04:45         &      &         &           &              & 20.7              &   -93.6         &             &  0.11\\ 
USG-03 & 1997-10-09    & 18:47:15         &      &         &           &              & 31.8              &  -106.0         &   37.0      &  0.53\\
USG-04 & 2000-01-18    & 08:33:58         &      &         &           &              & 24.3              &   -94.9         &             &  0.12\\
USG-05 & 2000-08-25    & 01:12:25         &      &         &           &              & 14.5              &  -106.1         &             &   3.1\\
USG-06 & 2005-11-15    & 05:19:07         &      &         &           &              & 26.3              &  -113.4         &   32.4      & 0.089\\
USG-07 & 2015-07-19    & 07:06:26         & 17.8 &   9.4   &  13.0     & 7.8          & 20.6              &   -87.6         &   22.0      & 0.082\\
USG-08 & 2019-02-01    & 18:17:10         & 16.3 &  -2.4   &  13.6     & 8.7          & 22.5              &   -83.8         &   23.7      &   1.4\\
USG-09 & 2019-06-22    & 21:25:48         & 14.9 & -13.4   &   6.0     & 2.5          & 14.9              &   -66.2         &   25.0      &   6.0\\
USG-10 & 2020-04-28    & 05:43:17         &      &         &           &              & 20.1              &  -109.4         &             & 0.076\\\hline  
  \end{tabular}
\end{table*}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{./figures/energies_boxplot}
  \caption{Comparison between released energies of bolides detected by the Geostationary Lightning Mapper and USG sensors.}
  \label{fig:boxplot}
\end{figure}

\subsection{GPS data}
\label{ssec:GPS}
%This material is based on services provided by the GAGE Facility, operated by UNAVCO, Inc., with support from the National Science Foundation and the National Aeronautics and Space Administration under NSF Cooperative Agreement EAR-1724794.

%We got RINEX data from 3 to 7 stations depending of the event location and data availability that surround the event place in all directions as possible. A list of the stations where we got RINEX data is available in table \ref{tab:table-stations}. Most of the stations lie in mexican territory, but in some cases we required data from other stations to cover events near the mexican frontier at north or south.

%\clearpage
%\onecolumn
%\footnotesize
%\begin{landscape}
%%\begin{table*}
%%    \centering
%  \begin{longtable}{llllllp{12cm}}
%      \caption{List of GPS stations used for this work.}
%      \label{tab:table-stations}
%      \endfirsthead
%      \endhead
%    \hline
%    Station name & Latitude & Longitude & \multicolumn{3}{c}{Events ID}  & Citation \\\hline
%    \multirow{4}{*}{BAR1\hyperlink{Hudnut}{${}^1$}\hyperlink{Hudnut2}{${}^5$}} & \multirow{4}{*}{33.48} & \multirow{4}{*}{-119.03} & \multirow{4}{*}{GLM-07} & \multirow{4}{*}{GLM-12} & \multirow{4}{*}{GLM-13} & UNAVCO Community, Hudnut, Kenneth, King, Nancy, Aspiotes, Aris G., Borsa, Adrian A., Determan, \\
%    &&&&&& Daniel N., Galetzka, John E., Stark, Keith F., 2005, SCIGN-PBO Nucleus GPS Network - BAR1-Santa \\
%    &&&&&& Barbara Island One P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \\
%    &&&&&& \url{https://doi.org/10.7283/T5668BHN}.\\\hline
